### 2024-03-27 doc/sections/mav-syntax.tex
Add AtomAxiom

Rule~(\ref{rule:ll-axiom}) is the axiom rule in the usual one-sided presentation of linear logic.

@@

In BV, the $\vParr$ connective plays the same role as the comma does in the antecedent of a linear logic sequent, and the $\vUnit$ plays the same role as the empty sequent, which would give us rule~(\ref{rule:bv-axiom-bad}) for BV.

@@

Therefore, the axiom for BV is actually rule~(\ref{rule:bv-axiom-good}), where $\vC$ is a one-hole structure context.
In BV, all inference rules are unary.
A branching inference rule enforces \emph{disjointness} of its premise derivations.

@@

  $\inlineequation[rule:ll-axiom]{%

@@

  $\inlineequation[rule:bv-axiom-bad]{%

@@

  $\inlineequation[rule:bv-axiom-good]{%

@@

  $\inlineequation[rule:mav-axiom]{%
      \vP\vParr\vDual\vP\vInferFrom\vUnit}$
  \\[1\baselineskip]
  $\inlineequation[rule:ll-cut]{%

@@

  $\inlineequation[rule:ll-cut]{%
          \vGG,\vGG',\vUnit,\vGD,\vGD'}{%

@@

  $\inlineequation[rule:bv-cut-good]{%

@@

  $\inlineequation[rule:mav-cut]{%
      \vP\vParr\vDual\vP\vInferFrom\vUnit}$

@@

The inference rules of MAV are presented as a \emph{rewrite rules} on structures.
We write $\vP\vInferFrom\vQ$ when $\vP$ can be inferred from $\vQ$.
The axiom of linear logic is typically given as \ref{rule:axiom-ll}.
In MAV, the structure connective $\vParr$ plays the same role as the comma does in the antecedent of a linear logic sequent.
  \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \vP\vParr\vDual\vP
     & \vInferFrom
     & \vUnit
     & \RuleLabel{axiom}
    \vUnit
     & \vInferFrom
     & \vP\vTens\vDual\vP
     & \RuleLabel{cut}
    \vUnit\vWith\vUnit
     & \vInferFrom
     & \vUnit
     & \RuleLabel{tidy}
    (\vP\vTens\vQ)\vParr\vR
     & \vInferFrom
     & \vP\vTens(\vQ\vParr\vR)
     & \RuleLabel{switch}
    (\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vParr\vR)\vSeq(\vQ\vParr\vS)
     & \RuleLabel{sequence}
    \vP\vPlus\vQ
     & \vInferFrom
     & \RuleLabel{left}
    \vP\vPlus\vQ
     & \vInferFrom
     & \vQ
     & \RuleLabel{right}
    (\vP\vWith\vQ)\vParr\vR
     & \vInferFrom
     & (\vP\vParr\vR)\vWith(\vQ\vParr\vR)
     & \RuleLabel{external}
    (\vP\vSeq\vQ)\vWith(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vWith\vR)\vSeq(\vQ\vWith\vS)
     & \RuleLabel{medial}
    \vC\vPlug\vP\vInferFrom\vC\vPlug\vQ
     & \text{if}
     & \vP\vInferFrom\vQ
     & \RuleLabel{mono}
    % \\
    % \vP\vInferFrom\vQ
    %  & \text{if}
    %  & \vP\vEquiv\vQ
    %  & \RuleLabel{equiv}

We write $\vInferFrom*$ for the reflexive, transitive closure of $\vInferFrom$.
We write $\vBiInfer$ for the symmetric core of $\vInferFrom$ (\ie ${\vBiInfer} = {\vInferFrom\cap\vInferTo}$) and $\vBiInfer*$ for the reflexive, transitive closure of $\vBiInfer$.

\begin{remark}
  The structural connectives of BV are usually presented as lists, distinguished only by their brackets: $\vP\vTens\vQ$ is written as $\vls(\vP;\vQ)$; $\vP\vParr\vQ$ is written as $\vls[\vP;\vQ]$; and $\vP\vSeq\vQ$ is written as $\vls<\vP;\vQ>$.
\end{remark}

A \emph{normal} derivation is a derivation where \cref{rule:cut} is not used and \cref{rule:axiom} is restricted to atoms.

\begin{displaymath}
  \begin{array}{l@{\;\vInferFrom\;}l@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \vNeg\va\vParr\vPos\va
     & \vUnit
     & \RuleLabel{Axiom}
  \end{array}
\end{displaymath}

\begin{itemize}
  \item Structural equivalences are explicit.
  \item Inference is a congruence.
  \item Inference rules are unary.
  \item Relate to usual presentation of MALL.
  \item Relate to usual presentation of BV.
\end{itemize}

>>>>

Rule~(\ref{rule:AxiomLL}), shown below, is the axiom rule in the usual one-sided presentation of linear logic.

@@

In BV, the $\vParr$ connective plays the same role as the comma does in the antecedent of a linear logic sequent, and the $\vUnit$ plays the same role as the empty sequent, which would give us rule~(\ref{rule:AxiomBV-bad}) for BV.

@@

Hence, the axiom for BV is rule~(\ref{rule:AxiomBV}), where $\vC$ is a one-hole structure context.

@@

  $\inlineequation[rule:AxiomLL]{%

@@

  $\inlineequation[rule:AxiomBV-bad]{%

@@

  $\inlineequation[rule:AxiomBV]{%

@@

  % \qquad
  % $\inlineequation[rule:mav-axiom]{%
  %     \vP\vParr\vDual\vP\vInferFrom\vUnit}$
Rule (\ref{rule:CutLL}) is the cut rule in the usual one-sided presentation of linear logic.
In rule (\ref{rule:CutLL}), as in any branching inference rule, the branching enforces the \emph{disjointness} of the premise derivations.
Hence, it plays the same role as branching does in sequent calculus.
This would give us rule (\ref{rule:CutBVBad}) for BV.
However, as BV's inference rules can work arbitrarily deep in the structure, the surrounding context of pars is unnecessary---and too restrictive. Hence, the cut for BV is rule (\ref{rule:CutBV}).
  $\inlineequation[rule:CutLL]{%

@@

  $\inlineequation[rule:CutBVBad]{%
          \vGG\vParr\vGG'\vParr\vGD\vParr\vGD'}{%
          \vlhy{%
            (\vGG\vParr\vP\vParr\vGG')
            \vTens
            (\vGD\vParr\vDual\vP\vParr\vGD')}}}}$

@@

  $\inlineequation[rule:CutBV]{%

@@

          \vC\vPlug{\vUnit}}{%
          \vlhy{\vC\vPlug{\vP\vTens\vDual\vP}}}}}$

@@

Beautifully, internalising branching makes the symmetry between the axiom and cut plain to see.

Proof trees are a cumbersome presentation for BV's derivations---they are convenient for trees, but cumbersome in the absence of branching.
\emph{Rewriting systems}, on the other hand, are a well-known and convenient notation for such derivations.
Hence, in MAV, inference rules are presented as rewrite rules.

\emph{Inference}, written $\vInferFrom$, is the smallest relation defined by the following axioms:
% NOTE: These shenanigans are for cross-text table alignment.
\newlength{\vInferFromDefnWidthOfLBL}%
\settowidth{\vInferFromDefnWidthOfLBL}{$\RuleName{AtomAxiom}$}%
\newlength{\vInferFromDefnWidthOfLHS}%
\settowidth{\vInferFromDefnWidthOfLHS}{$(\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)$}%
\newlength{\vInferFromDefnWidthOfRHS}%
\settowidth{\vInferFromDefnWidthOfRHS}{$(\vP\vParr\vR)\vWith(\vQ\vParr\vR)$}%
\begin{displaymath}
  \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \vP\vParr\vDual\vP
     & \vInferFrom
     & \vUnit
     & \makebox[\vInferFromDefnWidthOfLBL][l]{$\RuleLabel{Axiom}$}
    \\
    \vUnit
     & \vInferFrom
     & \vP\vTens\vDual\vP
     & \RuleLabel{Cut}
    \\
    \vUnit\vWith\vUnit
     & \vInferFrom
     & \vUnit
     & \RuleLabel{Tidy}
    \\
    (\vP\vTens\vQ)\vParr\vR
     & \vInferFrom
     & \vP\vTens(\vQ\vParr\vR)
     & \RuleLabel{Switch}
    \\
    (\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vParr\vR)\vSeq(\vQ\vParr\vS)
     & \RuleLabel{Sequence}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vP
     & \RuleLabel{Left}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vQ
     & \RuleLabel{Right}
    \\
    (\vP\vWith\vQ)\vParr\vR
     & \vInferFrom
     & (\vP\vParr\vR)\vWith(\vQ\vParr\vR)
     & \RuleLabel{External}
    \\
    (\vP\vSeq\vQ)\vWith(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vWith\vR)\vSeq(\vQ\vWith\vS)
     & \RuleLabel{Medial}
    \\
    \vC\vPlug\vP\vInferFrom\vC\vPlug\vQ
     & \text{if}
     & \vP\vInferFrom\vQ
     & \RuleLabel{Mono}
    % \\
    % \vP\vInferFrom\vQ
    %  & \text{if}
    %  & \vP\vEquiv\vQ
    %  & \RuleLabel{Equiv}
  \end{array}
\end{displaymath}
(If $\vP\vInferFrom\vQ$, we say that $\vP$ can be inferred from $\vQ$, \ie the arrow points from conclusion to premise.)

\emph{Derivation}, written $\vInferFrom*$ is the reflexive, transitive closure of inference.
\emph{Invertible inference}, written $\vBiInfer$, is the symmetric core of $\vInferFrom$, \ie ${\vBiInfer} = {\vInferFrom\cap\vInferTo}$.
\emph{Invertible derivation}, written $\vBiInfer*$, is the symmetric core of derivation, or, alternatively, the reflexive, transitive closure of invertible inference.

A derivation is in \emph{normal form} if it does not use \cref{rule:Cut} and  use of \cref{rule:Axiom} is restricted to atoms.
\begin{displaymath}
  \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \makebox[\vInferFromDefnWidthOfLHS][l]{$\vNeg\va\vParr\vPos\va$}
     & \vInferFrom
     & \makebox[\vInferFromDefnWidthOfRHS][l]{$\vUnit$}
     & \RuleLabel{AtomAxiom}
  \end{array}
\end{displaymath}
Usually, these restrictions correspond to the outcome cut elimination and identity expansion, respectively.
\begin{remark}
  The structural connectives of BV are usually presented as lists, distinguished only by their brackets: $\vP\vTens\vQ$ is written as $\vls(\vP;\vQ)$; $\vP\vParr\vQ$ is written as $\vls[\vP;\vQ]$; and $\vP\vSeq\vQ$ is written as $\vls<\vP;\vQ>$.
\end{remark}
